---
title: "Manipulating Time Series Data in R"
output: github_document
---

### Module 01 - What Is Time Series Data?

A **time series** is a collection of data points ordered sequentially over time.

Time series objects: work better with specialized tools!

```{r}
# Open data
maunaloa <- readRDS("maunaloa.Rds")

# Load ggplot2
library(forecast, ggplot2)

# Plot time serie object
autoplot(maunaloa)
```

Some of the most essential information about your dataset is its **summary statistics**, which include attributes like the mean, minimum, maximum, median, and so on.

```{r}
summary(maunaloa)
```

Date-Time classes:

- Date: "2022-08-01"

- POSIXct: "2022-08-09 20:17:00 UTC"

- **lubridate**::as_date() vs base::as.Date() > better support + warnings for invalids + easier conversions

```{r}
# Char vector
weekends <- c("2019-01-05", "2019-01-06", "2019-01-12", "2019-01-13", "2019-01-19")
class(weekends)

library(lubridate)

# Convert to date
weekends_date <- as_date(weekends)
class(weekends_date)
```

ISO 8601: year-month-day

To convert Date format: **lubridate::parse_date_time()**

```{r}
# Char vector
dates <- c("2019-01-01", "2019-01-02", "2019-01-03", "2019-01-04", "2019-01-05")
class(dates)

# Convert to Date
dates_converted <- as_date(dates)
class(dates_converted)

# Mixed char vector: / and -
dates2 <- c("2019-01-01", "2019-01-02", "01/03/2019", "2019-01-04", "2019-01-05")

# Convert different formats all at once!
parse_date_time(dates2, orders = c("%Y-%m-%d", "%m/%d/%Y"))
```

