---
title: "Introduction to R for Finance"
output: html_document
---

### Module 02 - Vectors and Matrices

**c()**ombine to create crate vectors

Note: a vector can only be composed of one data type!

```{r}
# Numeric
ibm_stock <- c(159.82, 160.02, 159.84)

# Strings
finance <- c("stocks", "bonds", "investments")

# Logical
logic <- c(TRUE, FALSE, TRUE)
```

Use functions **plot()** or **hist()** to visualize datasets.

```{r}
data("mtcars") # Built in dataset
v <- mtcars$mpg
hist(v, xlab="MPG", ylab="Frequency", main="Distribution")

```

Use function **sum()** to sum over the entire vector.

```{r, collapse=TRUE}
ret <- c(7, 9) # Return
weight <- c(.2, .8) # Weight
companies <- c("Microsoft", "Sony") # Company Name

# Assign company names to your vectors
names(ret) <- companies
names(weight) <- companies

# Multiply the returns and weights together 
ret_X_weight <- ret * weight

# Print ret_X_weight
ret_X_weight

# Sum to get the total portfolio return
portf_ret <- sum(ret_X_weight)

cat("Total Portfolio Return:", portf_ret)
```
You can subset a vector using brakets: [ ]

```{r, collapse=TRUE}
print("First five items:")
v[1:5]

print("All except first:")
v[-1]

# Last item of a vecotr
print("Last item:")
cat(v[length(v)], tail(v, n=1))
```
Create a Matrix using function **matrix()**

```{r}
# Vector
my_vector <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)

# 3x3 matrix
my_matrix <- matrix(data = my_vector, nrow = 3, ncol = 3)

my_matrix

# Filling across using byrow = TRUE
matrix(data = c(2, 3, 4, 5), nrow = 2, ncol = 2, byrow = TRUE)
```
Merge vectors to create a matrix using **cbind()** or **rbind()**

```{r, collapse=TRUE}
apple <- c(100, 101, 102)
ibm <- c(55, 45, 47)

cbind_stocks <- cbind(apple, ibm)
cbind_stocks

rbind_stocks <- rbind(apple, ibm)
rbind_stocks

plot(cbind_stocks)
```
Use function **cor()** to calculate correlation between variables

```{r}
cor(cbind_stocks)
```
To subset a matrix: **my_matrix[row, col]**

```{r, collapse=TRUE}
cbind_stocks[1,] # First row
cbind_stocks[,"apple"] # apple column
cbind_stocks[1,1]
```

