---
title: "Introduction to R for Finance"
output: github_document
---

### Module 03 - DataFrames

DataFrames can store different types of data.

Use constructor **data.frame()** to create one.

```{r}
# Variables
company <- c("A", "A", "A", "B", "B", "B", "B")
cash_flow <- c(1000, 4000, 550, 1500, 1100, 750, 6000)
year <- c(1, 3, 4, 1, 2, 4, 5)

# DataFrame
cash <- data.frame(company, cash_flow, year)
cash
```
```{r}
head(cash, n=4)

tail(cash, n=3)

# Structure of the DataFrame
str(cash)
```

Use **colnames()** and **rownames()** to rename col/row functions.

```{r}
colnames(cash)

colnames(cash) <- c("company", "cashflow", "year")
colnames(cash)

rownames(cash)
```
To subset a DataFrame use brackets: **dataframe[row, col]**

```{r}
cash[3,2]
cash[,"company"]
```
To acccess a column in a df, you can also use: **dataframe$col**

```{r}
cash$year

cash$cashflow * 2

# Delete a column
cash$company <- NULL
cash
```
To filter a df with conditions, use function **subset()**

```{r}
subset(cash, company == "B")

subset(cash, year == 1)
```

To append a new column, **dataframe$new_col**

```{r}
cash$double_year <- cash$year * 2
```

To append a new row, use **rbind()**

```{r}
df1 <- c(10, 1000)
names(df1) <- c("eta", "valore")

df2 <- c(15, 800)
names(df2) <- c("eta", "valore")

df <- rbind(df1, df2)
df
```
Exercise: calculate the present value of projected cash flows.

```{r}
# Present value of $4000, in 3 years, at 5%
present_value_4k <- 4000 * (1.05)^-3

# Present value of all cash flows at 5%
cash$present_value <- cash$cashflow * (1.05)^-cash$year
cash
```

