---
title: "Introduction to R for Finance"
output: github_document
---

### Module 04 - Factors

Factors are used to work with categorical variables.

```{r}
# Vector to transform
credit_rating <- c("BB", "AAA", "AA", "CCC", "AA", "AAA", "B", "BB")

# Factor constructor
credit_factor <- factor(credit_rating)
credit_factor

# Call str() on credit_rating
str(credit_rating)

# Call str() on credit_factor
str(credit_factor)
```

To access and/or rename the levels of a factor by **levels()**

```{r}
# Print levels
levels(credit_factor)

# Rename levels
levels(credit_factor) <- c("2A", "3A", "1B", "2B", "3C")
levels(credit_factor)
```

```{r}
plot(credit_factor)
```

```{r}
# Ranking
AAA_ranks <- seq(1, 100)

# Factor & groups
AAA_factor <- cut(AAA_ranks, breaks = c(0,25,50,75,100))

# Rename levels
levels(AAA_factor) <- c("low", "medium", "high", "very_high")

plot(AAA_factor)
```

For ordered factors:

- When creating a factor, specify ordered = TRUE and add unique levels in order from least to greatest.

- For an existing unordered factor like credit_factor, use the ordered() function and specify levels.

```{r}
# Vector
credit_rating

# Ordered
credit_factor_ordered <- factor(credit_rating, ordered = TRUE, levels = c("AAA", "AA", "B", "BB", "CCC"))

credit_factor_ordered

plot(credit_factor_ordered)
```

To subset Factors, use brackets []

```{r}
# Subsetting
credit_factor[1:3]
credit_factor[2]

# Remove items, without drop
keep_level <- credit_factor[c(-3,-5)]
plot(keep_level)

# Remove items, with drop = remove levels
drop_level <- credit_factor[c(-3,-5), drop=TRUE]
plot(drop_level)
```


