---
title: "Intermediate R for Finance"
output: github_document
---

### Module 01 - Dates

The two main classes of data for this are Date and POSIXct:

- Date is used for calendar date objects like "2015-01-22"

- POSIXct is a way to represent datetime objects like "2015-01-22 08:39:40 EST"

```{r}
# Current date
Sys.Date()

# Current date+time
Sys.time()

today <- Sys.Date()
class(today)
```

To convert string (yyyy-mm-dd) to Date use as.Date():

```{r}
# Date
crash <- as.Date("2008-09-29")
crash

# String as Numeric (= days since January, 1 1970)
as.numeric(crash)

as.numeric(Sys.time())
```

To convert multiple dates from character to date format, you can do this all at once using *vectors*.

```{r}
# Vector
dates <- c("2017-02-05", "2017-02-06", "2017-02-07", "2017-02-08")

# Names
names(dates) <- c("Sunday", "Monday", "Tuesday", "Wednesday")
dates

# Subset "Monday"
dates[2]
```

Specify Date format with its parameter:

```{r collapse=TRUE}
as.Date("Ago 30, 1930", format = "%b %d, %Y")

as.Date("12.05.00", format="%d.%m.%y")

# Vector
char_dates <- c("1gen17", "2gen17", "3gen17", "4gen17", "5gen7")

# Create dates using as.Date() and the correct format 
dates <- as.Date(char_dates, format = "%e%b%y")

# Use format() to go from "2017-01-04" -> "Jan 04, 17"
format(dates, format = "%b %d, %y")

# Use format() to go from "2017-01-04" -> "01,04,2017"
format(dates, format = "%m,%d,%Y")
```

Date difference:

```{r, collapse=TRUE}
today <- Sys.Date()
today

tomor <- Sys.Date() + 1
tomor

# Difference
tomor - today

# Difference format
difftime(tomor, today, units = "hours")
```

A few functions useful for extracting date components: **months() weekdays() quarters()**

```{r collapse=TRUE}
dates <- as.Date(c("2017-01-02", "2017-05-03", "2017-08-04", "2017-10-17"))

# Months
months(dates)

# Quarters
quarters(dates)

dates2 <- as.Date(c("2017-01-02", "2017-01-03", "2017-01-04", "2017-01-05"))

# Assign the weekdays() of dates2 as the names()
names(dates2) <- weekdays(dates2)

# Print dates2
dates2
```